Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-05-14 Chapter-3/3-TheLArIATExperiment.tex
s

Events satisfying the simplest scenario of one single hit in each of the four wire chambers form the ``Picky Track" sample.  We construct another, higher statistics sample, where we loosen the requirements on single hit and wire chamber efficiency: the ``High Yield" sample. For LArIAT Run II, the High Yield sample is about three times the Picky Tracks statistics.  For the first measurements of the LArIAT hadronic cross section, we use the Picky Tracks sample because the uncertainty on the momentum is smaller and the comparison with the beamline MC results is straightforward compared with the High Yield sample;  a possible future update and cross check of these analysis would be the use of the High Yield sample.

>>>>

Events satisfying the simplest scenario of one single hit in each of the four wire chambers form the ``Picky Track" sample.  We construct another, higher statistics sample, where we loosen the requirements on single hit and wire chamber efficiency: the ``High Yield" sample. For LArIAT Run II, the High Yield sample is about three times the Picky Tracks statistics.

### 2018-05-14 Chapter-4/4-Methodology.tex
s

\chapter{Total Ha($\pi^{-}$,Ar)dronic Cross Section Measurement Methodology}\label{ch:Interactions}
This chapter describes the general procedure employed to measure a total hadronic differential cross section in LArIAT.
Albeit with small differences, both the  ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section measurements rely on the same procedure described in details in the following sections. We start by selecting the particle of interest using a combination of beamline detectors and TPC information (Section \ref{ch:ParticleSelectionMeEquationthod}). We then perform a handshake between the beamline information and the TPC tracking to assure the selection of the right TPC track (Section \ref{ch:WC2TPCMatchMethod}). Finally, we apply the ``thin slice" method and measure the ``raw" hadronic cross section (Section \ref{ch:ThinSliceMethod}). A series of corrections are then evaluated to obtain the ``true" cross section (Section \ref{ch:MCCorrections}).

@@

\subsubsection{Selection of Beamline Events}\label{ch:beamlineDetectorsData}
As shown in equation \ref{eq:enFF}, we leverage the beamline particle identification and momentum measurement before entering the TPC as in input to evaluate the kinetic energy for the hadrons used in the  cross sections measurements. Thus, we select the LArIAT data to keep only events whose wire chamber and time of flight information is registered (line 2 in in Table \ref{tab:beamlineDataSelection}). Additionally, we perform a check of the plausibility of the trajectory inside the beamline detectors: given the position of the hits in the four wire chambers, we make sure the particle's trajectory does not cross any impenetrable material such as the collimator and the magnets steel (line 3 in in Table \ref{tab:beamlineDataSelection}).

\subsubsection{Particle Identification in the beamline}
In data, the main tool to establish the identity of the hadron of interest is the LArIAT tertiary beamline, in its function of mass spectrometer. We combine the measurement of the time of flight, $TOF$, and the beamline momentum, $p_{Beam}$, to reconstruct the invariant mass of the particles in the beamline, $m_{Beam}$, as follows
\begin{equation}
m_{Beam} = \frac{p_{Beam}}{c}\sqrt{\biggl(\frac{TOF*c}{l}\biggr)^2 -1},
\label{eq:mass}
\end{equation}
 where $c$ is the speed of light and $l$ is the length of the particle's trajectory between the time of flight paddels.

@@

\subsubsection{TPC Selection: Halo mitigation }\label{ch:pileUp}
The secondary beam impinging on LArIAT secondary target produces a plethora of particles which propagates downstream. The presence of upstream and downstream collimators greatly abates the number of particles tracing down the LArIAT tertiary beamline. However, it is possible that more than one particle sneaks into the LArTPC during its readout time: the TPC readout is triggered by the particle  firing the beamline detectors, but particles from the beam halo might be present in the TPC at the same time. We call ``pile up" the additional traces in the TPC. We adjusted the primary beam intensity between LArIAT Run I and Run II to reduce the presence of events with high pile up particles in the data sample. For the cross section analyses, we remove events with more than 4 tracks in the first 14 cm upstream portion of the TPC from the sample (line 6 in in Table \ref{tab:beamlineDataSelection}).

\subsubsection{TPC Selection: Shower Removal}\label{ch:electrons}
In the case of the ($\pi^-$,Ar) cross section, the resolution of  beamline mass spectrometer is not sufficient to select a beam of pure pions. In fact, muons and electrons survive the selection on the beamline mass. It is important to notice that the composition of the negative polarity beam is mostly pions, as will be discussed in section \ref{ch:beamlineComposition}.
Anyhow, we devise a selection on the TPC information to mitigate the presence of electrons in the sample used for the pion cross section. The selection relies on the different topologies of a pion and an electron event in the argon: while the former will trace a track inside the TPC active volume, the latter will tend to ``shower", i.e. interact with the medium, producing bremsstrahlung photons which pair convert into several short tracks. In order to remove the shower topology, we create a region of interest (ROI) around the TPC track corresponding to the beamline particle (more details on this in the next section). We look for short tracks contained in the ROI, as depicted in figure \ref{fig:showerFilt}:  if more then 5 tracks shorter than 10 cm are in the ROI, we reject the event. Line 8 in in Table \ref{tab:beamlineDataSelection}) shows the number of events surviving this selection.

>>>>

\chapter{Total Hadronic Cross Section Measurement Methodology}\label{ch:Interactions}
This chapter describes the general procedure employed to measure a total hadronic differential cross section in LArIAT.
Albeit with small differences, both the  ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section measurements rely on the same procedure described in details in the following sections. We start by selecting the particle of interest using a combination of beamline detectors and TPC information (Section \ref{ch:ParticleSelectionMeEquationthod}). We then perform a handshake between the beamline information and the TPC tracking to assure the selection of the right TPC track (Section \ref{ch:WC2TPCMatchMethod}). Finally, we apply the ``thin slice" method and measure the ``raw" hadronic cross section (Section \ref{ch:ThinSliceMethod}). A series of corrections are then evaluated to obtain the ``true" cross section (Section \ref{ch:MCCorrections}).

@@

\subsection{Selection of Beamline Events}\label{ch:beamlineDetectorsData}
As shown in equation \ref{eq:enFF}, we leverage the beamline particle identification and momentum measurement before entering the TPC as in input to evaluate the kinetic energy for the hadrons used in the  cross sections measurements. Thus, we select the LArIAT data to keep only events whose wire chamber and time of flight information is registered (line 2 in in Table \ref{tab:beamlineDataSelection}). Additionally, we perform a check of the plausibility of the trajectory inside the beamline detectors: given the position of the hits in the four wire chambers, we make sure the particle's trajectory does not cross any impenetrable material such as the collimator and the magnets steel (line 3 in in Table \ref{tab:beamlineDataSelection}).

\subsection{Particle Identification in the Beamline}
In data, the main tool to establish the identity of the hadron of interest is the LArIAT tertiary beamline, in its function of mass spectrometer. We combine the measurement of the time of flight, $TOF$, and the beamline momentum, $p_{Beam}$, to reconstruct the invariant mass of the particles in the beamline, $m_{Beam}$, as follows
\begin{equation}
m_{Beam} = \frac{p_{Beam}}{c}\sqrt{\biggl(\frac{TOF*c}{l}\biggr)^2 -1},
\label{eq:mass}
\end{equation}
 where $c$ is the speed of light and $l$ is the length of the particle's trajectory between the time of flight paddels.

@@

\subsection{TPC Selection: Halo Mitigation }\label{ch:pileUp}
The secondary beam impinging on LArIAT secondary target produces a plethora of particles which propagates downstream. The presence of upstream and downstream collimators greatly abates the number of particles tracing down the LArIAT tertiary beamline. However, it is possible that more than one particle sneaks into the LArTPC during its readout time: the TPC readout is triggered by the particle  firing the beamline detectors, but particles from the beam halo might be present in the TPC at the same time. We call ``pile up" the additional traces in the TPC. We adjusted the primary beam intensity between LArIAT Run I and Run II to reduce the presence of events with high pile up particles in the data sample. For the cross section analyses, we remove events with more than 4 tracks in the first 14 cm upstream portion of the TPC from the sample (line 6 in in Table \ref{tab:beamlineDataSelection}).

\subsection{TPC Selection: Shower Removal}\label{ch:electrons}
In the case of the ($\pi^-$,Ar) cross section, the resolution of  beamline mass spectrometer is not sufficient to select a beam of pure pions. In fact, muons and electrons survive the selection on the beamline mass. It is important to notice that the composition of the negative polarity beam is mostly pions, as will be discussed in section \ref{ch:beamlineComposition}.
Anyhow, we devise a selection on the TPC information to mitigate the presence of electrons in the sample used for the pion cross section. The selection relies on the different topologies of a pion and an electron event in the argon: while the former will trace a track inside the TPC active volume, the latter will tend to ``shower", i.e. interact with the medium, producing bremsstrahlung photons which pair convert into several short tracks. In order to remove the shower topology, we create a region of interest (ROI) around the TPC track corresponding to the beamline particle (more details on this in the next section). We look for short tracks contained in the ROI, as depicted in figure \ref{fig:showerFilt}:  if more then 5 tracks shorter than 10 cm are in the ROI, we reject the event. Line 8 in in Table \ref{tab:beamlineDataSelection}) shows the number of events surviving this selection.

### 2018-05-14 Chapter-5/5-Preparatory.tex
s

\chapter{Preparatory Work}\label{ch:samples}
This chapter describes the preparatory work done on the the data and Monte Carlo samples used for the cross section analyses. This entails:

\begin{enumerate}
\item the MC production,
\item the energy calibration of the detector both in data and MC,
\item the optimization of the tracking algorithm for the total cross section analyses.
\end{enumerate}

\section{Construction of a Monte Carlo Simulation for LArIAT}
\subsection{G4Beamline}\label{ch:beamlineComposition}
\subsection{Data Driven MC}\label{sec:DDMC}

\section{Tracking Studies}

\section{Energy Calibration and Studies}

\section{Estimate of Energy Loss before the TPC}

\begin{comment}
\section{Construction of a Monte Carlo Simulation for LArIAT}
For the simulation of LArIAT events and their particle make up, we use a combination of two MC generators: the G4Beamline Monte Carlo and the Data Driven single particle Monte Carlo (DDMC). We use the G4Beamline MC to simulate the particle transportation in the beamline and calculate the particle composition of the beam just past the fourth Wire Chamber (WC4). In order to simulate the beam line particles after WC4 and in the TPC, we use the DDMC.
\subsection{G4Beamline}\label{beamlineComposition}
G4Beamline simulates the beam collision with the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors and the action of the LArIAT magnets, effectively accounting for particle transportation through the beam line from the LArIAT target until ``Big Disk", a fictional, void detector located just before the cryostat. 
 At the moment of this writing, G4Beamline does not simulated the responses of the beam line detectors. It is possible to interrogate the truth level information of the simulated particles in several points of the geometry. In order to ease the handshake between G4Beamline and the DDMC, we ask for the beam composition just after WC4.
Since LArIAT data are taken under different beam conditions, G4Beamline simulates separately the beam composition according to the magnets' settings and the secondary beam intensity. For the pion cross section analysis the relevant beam conditions are  secondary beam energy of 64 GeV, negative polarity magnet with current of 100 A and 60 A. For the kaon cross section analysis the relevant beam conditions is a secondary beam energy of 64 GeV, positive polarity magnet with current of 100 A.

>>>>

\chapter{Preparatory Work}\label{ch:samples}
%{\raggedleft ``\emph{Il dolce non lo mangi mai, ma qualche volta ti rifai.} \par}
%{\raggedleft \emph{Abbracciami}"\par}
%{\raggedleft -- Pietro Ciampi, L'amore e' tutto qui, 1971 -- \par}
%\vspace{0.5cm}

%{\raggedleft ``\emph{You never eat dessert, but, sometimes, you make up for it.} \par}
%{\raggedleft \emph{Hug me}"\par}
%{\raggedleft -- Pietro Ciampi, L'amore e' tutto qui, 1971 -- \par}
%\vspace{0.5cm}

This chapter describes the preparatory work done on the the data and Monte Carlo samples used for the cross section analyses. 
This entails the choice of the data set and the production of the information needed to construct the Monte Carlo Simulation~(section \ref{sec:dataSet}),  the construction and use of said Monte Carlo simulation~(section \ref{sec:MCSet}), the study and optimization of the tracking in the TPC for the cross section analyses~(section \ref{sec:TrackingStudies}),  the calibration of the calorimetry response and related energy studies~(section \ref{ch:energyCal}).

\section{Cross Section Analyses Data Set}\label{sec:dataSet}
We choose LArIAT Run-II as the data period for the  ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section analyses. 
Data taking for the this period started on 03/15/2016  and ended on 07/31/2016. 
Since we are interested in beamline and TPC information, we ask basic requirements on the operational status of the time of fight, wire chambers and TPC to form the good run list for this period, which we informally call ``lovely runs".

The subset of lovely runs  chosen for the  ($\pi^{-}$,Ar) total hadronic cross section analysis includes only the -60A and -100A magnet configurations in negative polarity, even if LArIAT explored several other beamline configurations during Run-II. The -60A and -100A combined data set accounts for approximately 90\% of the total Run-II negative polarity runs.   Since the production of beamline Monte Carlo depends on the wanted beamline configuration, the choice of only two beamline settings limits the need for beamline MC production.

Similarly, the subset of lovely runs chosen  for the (K$^{+}$,Ar)  total hadronic cross section analysis includes only the +60A and +100A magnet configurations in positive polarity. It should be noted that kaons are extremely rare in the +60A sample, thus the data sample for the (K$^{+}$,Ar) cross section after the mass selection is about 90\% +100A runs, as shown in Table \ref{tab:databreakdown}.

For the first measurements in LArIAT that uses both beamline and TPC information, we choose strict requirements on the reconstruction of the WC tracks, the so-called ``Picky Track" sample (see \ref{sec:MWPCfunc}). This choice presents two advantages:  the uncertainty on the momentum reconstruction for the ``Picky Tracks" sample is smaller compared to the ``High Yield" sample, and the comparison with the beamline MC results is straightforward. A possible future update and cross check of these analysis would be the use of the High Yield sample, where the statistics is about three times higher.

The breakdown of beamline events as a function of the magnets settings is shown in Table \ref{tab:databreakdown}. 
The choice of the data sets determines the production of beamline MC and serves as basis for the production of Data Driven MC, as shown in the next sections.

\begin{table}[b]
\centering
\begin{tabular}{|l|c|c|c|}  
\hline
                                                              & I = 60 A          & I = 100 A   & Total     \\ \hline
Data Events after $\pi/\mu/e$ Mass Selection     &     67068          &  71413  & 138481 \\ \hline
Data Events after $K$ Mass Selection                &     274              &   2563   & 2837  \\ \hline
\end{tabular}
\caption{Number of data events which fit the $\pi/\mu/e$ or $K$ mass hypothesis as a function of magnet settings.}
\label{tab:databreakdown}
\end{table}

\section{Construction of a Monte Carlo Simulation for LArIAT}\label{sec:MCSet}
For the simulation of LArIAT events and their particle make up, we use a combination of two MC generators: the G4Beamline Monte Carlo and the Data Driven single particle Monte Carlo (DDMC). We use the G4Beamline MC to simulate the particle transportation in the beamline and calculate the particle composition of the beam just after the fourth Wire Chamber (WC4). In order to simulate the beamline particles after WC4 and in the TPC, we use the DDMC.

\subsection{G4Beamline}\label{ch:beamlineComposition}
G4Beamline simulates the beam collision with the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors and the action of the LArIAT magnets, effectively accounting for particle transportation through the beam line from the LArIAT target until ``Big Disk", a fictional, void detector located just before the LArIAT cryostat. 
 At the moment of this writing, G4Beamline does not simulated the responses of the beam line detectors. It is possible to interrogate the truth level information of the simulated particles in several points of the geometry. In order to ease the handshake between G4Beamline and the DDMC, we ask for the beam composition just after WC4.
Since LArIAT data are taken under different beam conditions, G4Beamline simulates separately the beam composition according to the magnets' settings and the secondary beam intensity. For the pion cross section analysis the relevant beam conditions are  secondary beam energy of 64 GeV, negative polarity magnet with current of 100 A and 60 A. For the kaon cross section analysis the relevant beam conditions is a secondary beam energy of 64 GeV, positive polarity magnet with current of 100 A.

\subsection{Data Driven MC}\label{sec:DDMC}
\subsection{Estimate of Energy Loss before the TPC}

\section{Tracking Studies}\label{sec:TrackingStudies}
\subsection{Study of WC to TPC Match}\label{ch:WC2TPCMatchOptimization}

\section{Energy Calibration and Studies}\label{ch:energyCal}

\begin{comment}
\section{Construction of a Monte Carlo Simulation for LArIAT}
\subsection{G4Beamline}\label{beamlineComposition}
